Treatment heterogeneity in comparative effectiveness of teriparatide vs bisphosphonates: multi-database cohort study

### 2021-12-06 submission/manuscript.rmd
add risk stratified results and update methods

##     Statistical analyses
We carried out two sets of analyses. First, we derived overall
treatment effect estimates of teriparatide compared to oral
bisphosphonates regarding the three outcomes of interest. To account
for potential measured confounding we developed separate large-scale
propensity score models within each database based on LASSO logistic
regression using the same predefined set of measured covariates
[@Tian2018]. We estimated treatment effects using Cox proportional
hazards models with treatment as the sole covariate fitted in the 1:4
propensity score-matched subset of the considered patient
population. Hazard ratios (HR) derived in different databases were
summarized using random effects meta-analysis.

For the second set of analyses, we used the standardized framework for
risk-based assessment of treatment effect heterogeneity, that follows
the directives of the PATH statement [@Kent2019; @KentEnE2019].
Following data extraction, the framework was applied in
two steps. First, we derived individualized risk predictions for the
three efficacy outcomes. We built the prediction models using LASSO
logistic regression on the propensity score matched (1:4)
subpopulation of the pooled treatment arms, aiming to remove any
effect of measured confounding that may cause differential fit of our
models across treatment arms. We considered the same large set of
candidate covariates as for the development of the propensity score
models.

For each outcome we used the derived prediction models to divide the
population in 4 equally-sized subgroups, based on the quarters of the
predicted risk distribution. Within each of these risk-based subgroups
we developed a new propensity score model. Our analyses were performed
on the propensity score matched (1:4) subset of the risk subgroup
subset. We derived relative effect estimates using Cox proportional
hazards models only with treatment as a predictor. Absolute effect
estimates were calculated based on the difference of the Kaplan-Meier
estimates, on day 730 after treatment initiation.

For each outcome we used the derived prediction models to divide the
population in 4 equally-sized subgroups, based on the quarters of the
predicted risk distribution. Within each of these risk-based subgroups
we developed a new propensity score model. Our analyses were performed
on the propensity score matched (1:4) subset of the risk subgroup
subset. We derived relative effect estimates using Cox proportional
hazards models only with treatment as a predictor. Absolute effect
estimates were calculated based on the difference of the Kaplan-Meier
estimates, on day 730 after treatment initiation.

@@

@@

\newpage
# Tables and figures

@@

\newpage
```{r overallPsDensity, cache = FALSE, echo = FALSE, warning=FALSE, fig.cap="Calibrated overall results"}
grid::grid.raster(tiff::readTIFF(here("figures/OverallPsDensity.tiff")))
```

\newpage
```{r overallNcPlot, cache = FALSE, echo = FALSE, warning=FALSE, fig.cap="Overall negative control plot."}
grid::grid.raster(tiff::readTIFF(here("figures/OverallNcPlot.tiff")))
```

\newpage
```{r plotMeta, echo=FALSE, warning=FALSE, fig.cap="Calibrated hazard ratios for the 3 outcomes of interest across the 3 considered databases using a set of 126 negative controls. Values below 1 favour teriparatide, while values above 1 favour bisphosphonates.", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here::here("figures/plotMeta.tiff")))
```

>>>>

##     Statistical analyses
We carried out two sets of analyses. First, we derived overall
treatment effect estimates of teriparatide compared to oral
bisphosphonates regarding the three outcomes of interest. To account
for potential measured confounding we developed separate large-scale
propensity score models within each database based on LASSO logistic
regression using the same predefined set of measured covariates
[@Tian2018]. To assume that treatment and comaparator cohorts stand in
equipoise we required that the majority of the patients in both have
preference scores between 0.3 and 0.7. We assumed that covariate
balance had been reached, if all after-adjustment baseline
characteristics had standardized absolute mean differences of less
than 0.1.  We estimated treatment effects using Cox proportional
hazards models with treatment as the sole covariate. To account for
observed confounding we used variable ratio matching (1:10) on the
propensity score using a caliper of 0.2 on the logit scale.  Hazard
ratios (HR) derived in different databases were summarized using
random effects meta-analysis.

For the second set of analyses, we used the standardized framework for
risk-based assessment of treatment effect heterogeneity, that follows
the directives of the PATH statement [@Kent2019; @KentEnE2019].
Following data extraction, the framework was applied in
two steps. First, we derived individualized risk predictions for the
three efficacy outcomes. We built the prediction models using LASSO
logistic regression on the propensity score matched (1:10)
subpopulation of the pooled treatment arms, aiming to remove any
effect of measured confounding that may cause differential fit of our
models across treatment arms. We considered the same large set of
candidate covariates as for the development of the propensity score
models.

For each outcome we used the derived prediction models to divide the
population in 4 equally-sized subgroups, based on the quarters of the
predicted risk distribution. Within each of these risk-based subgroups
we developed a new propensity score model. Our analyses were performed
on the propensity score matched (1:4) subset of the risk subgroup
subset. We derived relative effect estimates using Cox proportional
hazards models only with treatment as a predictor. Absolute effect
estimates were calculated based on the difference of the Kaplan-Meier
estimates, on day 365 after treatment initiation.

For each outcome we used the derived prediction models to stratify the
study population using two separate stratifcation schemes. In the
first approach, we divide the population into the lowest 75\% and the
top 25\% hip fracture risk subgroups. In the second approach, the
population is divided using existing guidelines on teriparatide
treatment based on major fracture risk (REF!!). Within each of these
risk-based subgroups we developed a new propensity score model. Our
analyses were performed on the propensity score matched (1:10) subset
of the risk subgroup subset. We derived relative effect estimates
using Cox proportional hazards models only with treatment as a
predictor. Absolute effect estimates were calculated based on the
difference of the Kaplan-Meier estimates, on day 365 after treatment
initiation.

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE}
metaCalibrateOverall <- readRDS(here("data/processed/metaCalibrateOverall.rds"))

suffix <- "age_50_tr_1_q_25_75"

calibrateOverallResults <- readRDS(
    here(
      file.path(
          "data/processed",
          paste0(
              paste(
                  "calibrateOverallResults",
                  suffix,
                  sep = "_"
              ),
              ".rds"
          )
      )
    )
)

```

In all databases adequate equipoise of the preference score
distributions was achieved (Figure \ref{fig:overallPsDensity}). After
matching all baseline covariates (>35,000 in each database) were well
balanced (Figure \ref{fig:overallCovariateBalance}). This indicates
that we were able account for the majority of the observed
condfounding present in all databases. However, negative control
analyses time falsely favored oral bisphosphonates over teriparatide,
which suggests the presence of residual unobserved confoundng (Figure
\ref{fig:overallNcPlot}).

In terms of hip fracture risk we estimated calibrated hazard ratios of
`r calibrateOverallResults %>% filter(database == "ccae", outcome == 101) %>% pull(hr)`
(`r calibrateOverallResults %>% filter(database == "ccae", outcome == 101) %>% pull(lower)`
to
`r calibrateOverallResults %>% filter(database == "ccae", outcome == 101) %>% pull(upper)`;
95\% CI), 
`r calibrateOverallResults %>% filter(database == "optum_extended_dod", outcome == 101) %>% pull(hr)`
(`r calibrateOverallResults %>% filter(database == "optum_extended_dod", outcome == 101) %>% pull(lower)`
to
`r calibrateOverallResults %>% filter(database == "optum_extended_dod", outcome == 101) %>% pull(upper)`;
95\% CI) and
`r calibrateOverallResults %>% filter(database == "optum_ehr", outcome == 101) %>% pull(hr)`
(`r calibrateOverallResults %>% filter(database == "optum_ehr", outcome == 101) %>% pull(lower)`
to
`r calibrateOverallResults %>% filter(database == "optum_ehr", outcome == 101) %>% pull(upper)`;
95\% CI), in CCAE, Optum-DOD and Optum-EHR respectively. The
calibrated meta-analytic hazard ratio of the overall effect of
teriparatide compared to oral bisphosphonates was
`r metaCalibrateOverall %>% filter(outcome == 101) %>% pull(hr)`
(`r metaCalibrateOverall %>% filter(outcome == 101) %>% pull(lower)`
to
`r metaCalibrateOverall %>% filter(outcome == 101) %>% pull(upper)`;
95\% CI). Results for major osteoporotic and vertebral fractures are
provided in the supplement.

When using the internally developed hip fracture risk prediction models to divide the population into lower 75\% and top 25\% risk subgroups, we observed....

@@

\newpage
# Tables and figures
\newpage
```{r overallPsDensity, cache = TRUE, echo = FALSE, warning=FALSE, fig.cap="Calibrated overall results"}
grid::grid.raster(tiff::readTIFF(here("figures/OverallPsDensity.tiff")))
```

@@

\newpage
```{r overallNcPlot, cache = TRUE, echo = FALSE, warning=FALSE, fig.cap="Overall negative control plot."}
grid::grid.raster(tiff::readTIFF(here("figures/OverallNcPlot.tiff")))
```

\newpage
```{r plotMeta, cache=TRUE, echo=FALSE, warning=FALSE, fig.cap="Calibrated hazard ratios for the 3 outcomes of interest across the 3 considered databases using a set of 126 negative controls. Values below 1 favour teriparatide, while values above 1 favour bisphosphonates.", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here::here("figures/plotMeta.tiff")))
```

\newpage
```{r plotMetaRiskStratified, cache=TRUE, echo=FALSE, warning=FALSE, fig.cap="Calibrated hazard ratios for the 3 outcomes of interest across the 3 considered databases using a set of 126 negative controls. Values below 1 favour teriparatide, while values above 1 favour bisphosphonates.", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here::here("figures/plotMetaRiskStratified.tiff")))
```

\newpage
```{r plotAbsoluteRiskStratified, cache=FALSE, echo=FALSE, warning=FALSE, fig.cap="Calibrated hazard ratios for the 3 outcomes of interest across the 3 considered databases using a set of 126 negative controls. Values below 1 favour teriparatide, while values above 1 favour bisphosphonates.", out.width = '100%'}
grid::grid.raster(tiff::readTIFF(here::here("figures/plotAbsoluteHip.tiff")))
```
